---
title: "Note10_1"
format: html
editor: visual
execute: 
  warning: false
---

## Regression 회귀분석

-   실증경제학에서는 꽃

-   Machine Learning (regression)

-   Recommendation. Logistic regression. Classfication(분류)

-   Deep Learning

-   Large Language Model (LLM). Chat GP**T** (transformer)

-   회귀모형regression model

$$
y_i = \beta_0 + \beta_1 x_i + \epsilon_i
$$

-   $y$ : 피설명변수, explained variable

-   $x$: 설명변수, explanatory variable

-   $\epsilon$ : 오차, error term

-   우리의 관심 $\beta_1 = 0$ 에 있는 거지.

-   estimator 추정량.

-   과녁. 총.

-   OLS (ordinary least square estimation)

-   MLE (maximum likelihood estimation)

-   기타 등등.

## OLS ordinary least square

-   least square 최소자승(제곱) 추정법

-   sum of squared errors (SSE) 최소

$$
\sum_{i} \epsilon_i^2
$$

-   Mincerian wage equation (민서 타입의 임금 방정식)

$$
\ln(w_i) = \beta_0 + \beta_1 edu_i + \beta_2 exp_i + e_i
$$

-   exp: 통제 control variable

## Simple regression 단순회귀

-   1개의 regressor (x)

-   데이터 로딩 "ceosal1.csv"

```{r}
rm(list = ls())
library("tidyverse")

readr::read_csv("ceosal1.csv", col_names = TRUE, show_col_types = FALSE) -> ceosal

# str(ceosal)
```

-   x 축에다 roe (return on equity), 수익률

-   y 축에다 salary (CEO의 연봉)으로 하는 산포도

```{r}
ceosal %>% 
  ggplot(aes(x = roe, y = salary)) +
  geom_point() +
  geom_smooth(method = "lm")
```

-   positive correlation between roe and salary

-   Is it statistically significant? 통계적으로 유의한가? 회구분석.

-   $$
    salary_i = a + b \times roe_i + e_i
    $$

-   $b=0$ 인지를 test 검정하는 것임.

-   귀무가설. the null-hypothesis

```{r}
lm.roe <- lm(salary ~ roe, data = ceosal)

summary(lm.roe)
```

```{r}
library("stargazer")

stargazer(lm.roe, 
          type = "text",
          keep.stat = c("n","rsq"))


```
